---
title: "Exploration"
author: "Aidan O'Hara"
date: "2023-05-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Diekmann2007

### digits

### Looking out for number one

```{r}
library(ISLR)
library(stringr)
library(dplyr)

magicNumber <- 54.2
otto <- Auto %>% select(-name) %>%
  mutate(firstDigitDisp = as.numeric(str_extract(displacement, "^\\d{1}"))) %>%
  mutate(dispMagic = displacement*magicNumber,
         firstDigitMagic = as.numeric(str_extract(dispMagic, "^\\d{1}"))) # see if scale invarience holds for car displacement
hist(otto$firstDigitDisp)
hist(otto$firstDigitMagic)

```

A skewed distribution of leading digits, aha! 
Is it along a log_10((n+1)/n) relationship?

Benford intially trialed his research on "baseball statistics, areas of river catchment, and ... addresses ..."

The next guy, Pinkham started messing with transformation tests. Currency and digit conversion. Scale Invarience.

Article goes on to measure digit frequency compared to predicted, the combination of all three data sets even more closely approximates Benfords.

### Exceptions
The lottery

Supposedly, and I don't buy this really, constrained measurements don't obey the digit frequency, I might argue they do obey their own digit frequency, which may just be a tranformed benford...

### Fraud
Apparentley, benfords law is not obeyed when people imagine numbers.

### Deriving Benford's law

```{r}
library(extraDistr)

# construct predicted frequencies.
start = c(1:9)
transform <- log10((start + 1)/start)
x <- rcat(1e5, transform)
plot(prop.table(table(x)), type = "h")
lines(1:9, dcat(1:9, transform), col = "red")

```

### The theory and application of Benford...

### derivation of benford