---
title: "Exploration"
author: "Aidan O'Hara"
date: "2023-05-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(ISLR)
library(stringr)
library(dplyr)

```

### Diekmann2007
Not the first Digit! Using Benford's Law to Detect Fraudulent Scientific Data




### digits

N! is a benford sequence
```{r}
a <- c(1:500)
b <- factorial(a)
c <- data.frame(a,b)
c <- c %>% mutate(firstDigit = as.numeric(str_extract(b, "^\\d{1}")))
hist(c$firstDigit)
```

### Looking out for number one

```{r}
magicNumber <- 54.2
otto <- Auto %>% select(-name) %>%
  mutate(firstDigitDisp = as.numeric(str_extract(displacement, "^\\d{1}"))) %>%
  mutate(dispMagic = displacement*magicNumber,
         firstDigitMagic = as.numeric(str_extract(dispMagic, "^\\d{1}"))) # see if scale invarience holds for car displacement
hist(otto$firstDigitDisp)
hist(otto$firstDigitMagic)

```

A skewed distribution of leading digits, aha! Is it along a log_10((n+1)/n) relationship?

Benford initially trialed his research on "baseball statistics, areas of river catchment, and ... addresses ..."

The next guy, Pinkham started messing with transformation tests. Currency and digit conversion. Scale Variance.

Article goes on to measure digit frequency compared to predicted, the combination of all three data sets even more closely approximates Benfords.

### Exceptions

The lottery

Supposedly, and I don't buy this really, constrained measurements don't obey the digit frequency, I might argue they do obey their own digit frequency, which may just be a tranformed benford...

Numbers that never have 1 as starting digit, numbers that always have 1 as starting digit, other digits may demonstrate? \### Fraud Apparentley, benfords law is not obeyed when people imagine numbers.

### Deriving Benford's law

```{r}
library(extraDistr)

# construct predicted frequencies.
start = c(1:9)
transform <- log10((start + 1)/start)
x <- rcat(1e5, transform)
plot(prop.table(table(x)), type = "h")
lines(1:9, dcat(1:9, transform), col = "red")

```

### The theory and application of Benford...

### derivation of benford, a vindication of Newcomb
Cool examples of sequences that obey BL, lots of lots of math for unpacking


# Examples

### Accounting data

### Elections data

### What about base 8?

### Restricted sets

### Likert scale test

Uniform dist \> 5 point likert \> sum of questions \> benford?

### A functional operation, for another day

